Key Provisioning for Internet of Things via Light

### 2015-09-10 paper/paper.tex
final fixes

@@

The Internet of Things (IoT) is becoming more and more of a reality every day.
It consists of small devices (called \textit{motes}), which are connected to the public Internet, \cite{atzori2010internet}.
They work together, allowing monitoring and controlling of several kinds of real-world systems.
An example are motes collecting information about parking lot occupancy allowing car drivers to find free spaces using their smart phones\footnote{see http://www.streetline.com/}.

@@

However, this comes at a price: by being accessible from the Internet, the attack surface of such systems is much larger than that of traditional embedded systems.
New, adapted security measurements had to be developed and implemented to react to this.
For example, \textit{Adaptive Key Establishment Scheme (AKES)} provides adaptable security measures for being used with the IEEE~802.15.4 network stack, which is commonly used by IoT devices, \cite{krentz15akes}.
Nevertheless, one problem remains: IoT setups usually consist of many nodes which all need to be initialized with keying material and parameters.
This is a major problem, because typical motes lack input/output devices which are securely usable without configured networking.

@@

In order to make the IoT reality, a specialized network stack is needed.
Motes have special constraints regarding energy and memory resources, e.g. batteries should work for multiple years and a mote's program memory will be very small\footnote{e.g. 48kb program flash for telosB mote, \cite{telosb}}.

@@

The target of the initialization phase is to tell the mote when the actual data starts.
The communication between the mote and the phone is one-way.
That means, the mote cannot communicate with the Android application.
Therefore, the application has to guess when the mote has successfully done calibration and synchronization.
This can either be done by using a fixed time (e.g. the application stays in calibration and synchronization phase for 10 seconds), or by user interaction (the user presses a button, when the mote indicates that it is ready via an LED).

@@

All data that is being sent is encoded using a \textit{Hamming(8,4)-code} so that two-bit errors can be detected and one-bit errors can be corrected.

The \textit{CRC32} is chosen as the checksum algorithm for the validation phase to avoid saving erroneous key material.
This could be caused by 3-or-more-bit errors that the Hamming-code failed to detect.

@@

The \textit{light process} implements the receiving side of the communication protocol.
A real-time timer called \textit{rtimer}, which is offered by the Contiki operating system, eliminates timer inaccuracies on the mote side.
The mote uses its light sensor to sense the current brightness.
Once the process is started, an LED-countdown (red $\rightarrow$ blue $\rightarrow$ green) indicates the start of the key initialization procedure.
A \mbox{\textit{2-means}} algorithm is used to create two clusters of \textit{bright} and \textit{dark} values and to compute their mean values~(\ref{sub:calibration}).
Subsequently a new value is classified as \textit{0} or \textit{1} depending on which mean value is closer to it.
After the successful calibration, synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the failure and the process should be restarted.
A green LED shows that the key material has been received successfully and was saved to the permanent flash memory of the mote.

@@

Another subject of debate is the \textit{Hamming(8,4)-code}.
While it is able to correct one-bit-errors and detect two-bit-errors, it also doubles the amount of data to be sent.
When analyzing the number of corrected bits depending on the transmission rate, the following can be shown:
With long period lengths, the Hamming code does not detect or correct any errors, because the transmission happens without any errors.
When accelerating the transmission, there is a point where many bits are corrected, but the transmission normally succeeds.
If the transmission speed is then set even higher, the whole transmission fails most of the time due to irrecoverable errors.
The borders are 15ms for the \textit{A3} and 63ms for the \textit{S2} (see Figure~\ref{fig:successful_transmissions}).
When we want to achieve the transmission speed of the \textit{A3} with 15ms, we can either use the 15ms and Hamming encoding to correct the bits, or we can halve the amount of data by deactivating the Hamming encoding and transmit with 31ms per bit\footnote{The values are rounded, that is why the half of 31 is 15.}.
So the transmission takes the same time, but actually increases the percentage of successful transmissions.
Errors can still be detected afterwards using the \textit{CRC32} checksum.
Therefore, the \textit{Hamming(8,4)-code} is not useful in that case.

@@

To simplify the initialization of several motes at once, two things need to be done.
First, the Android application should stay in calibration and synchronization phase, until the user presses a button.
Second, the mote should indicate via an LED that it reached the initialization phase~(\ref{sub:initialization}).
With those two concepts, the key transmission to several motes could work as follows:
The user lays out all motes in a grid next to each other.
Then the phone is placed on top of them and the app is started.
All motes are placed, so that they receive the flickering of the flash light.
The motes are started one after another.
Once the motes are running, the user waits for all of them to indicate via an LED, that they reached the initialization phase.
They are now all waiting for the initialization pattern.
The user presses a button in the application and the phone sends the initialization pattern followed by the data packet.
When the transmission is over, the user can see via a red LED which motes failed to receive the data correctly.
The process needs to be restarted with those motes.
However, this is not implemented and tested yet and could be part of future projects.

>>>>

\hyphenation{Ham-ming}

@@

The Internet of Things (IoT) is becoming more and more of a reality every day.
It consists of small devices (called \textit{motes}), which are connected to the public Internet, \cite{atzori2010internet}.
They work together, allowing monitoring and controlling of several kinds of real-world systems.
An example are motes collecting information about parking lot occupancy, allowing car drivers to find free spaces using their smart phones\footnote{see http://www.streetline.com/}.

@@

However, this comes at a price: by being accessible from the Internet, the attack surface of such systems is much larger than that of traditional embedded systems.
New security measures had to be developed and implemented to react to this.
For example, \textit{Adaptive Key Establishment Scheme (AKES)} provides adaptable security measures for being used with the IEEE~802.15.4 network stack, which is commonly used by IoT devices, \cite{krentz15akes}.
Nevertheless, one problem remains: IoT setups usually consist of many nodes which all need to be initialized with keying material and parameters.
This is a major problem, because typical motes lack input/output devices which are securely usable without configured networking.

@@

In order to make the IoT reality, a specialized network stack is needed.
Motes have special constraints regarding energy and memory resources, e.g. batteries should work for multiple years and a motes program memory will be very small\footnote{e.g. 48kb program flash for telosB mote, \cite{telosb}}.

@@

The target of the initialization phase is to tell the mote when the actual data transmission starts.
The communication between the mote and the phone is one-way.
That means, the mote cannot communicate with the Android application.
Therefore, the application has to guess when the mote has successfully done calibration and synchronization.
This can either be done by using a fixed time (e.g. the application stays in calibration and synchronization phase for 10 seconds), or by user interaction (the user presses a button, when the mote indicates that it is ready via an LED).

@@

All data that is being sent is encoded with a \textit{Hamming(8,4)-code} so that two-bit errors can be detected and one-bit errors can be corrected.

The \textit{CRC32} is chosen as the checksum algorithm for the validation phase to avoid saving erroneous key material.
This could be caused by 3-or-more-bit errors that the Hamming code failed to detect.

@@

The \textit{light process} implements the receiving side of the communication protocol.
A real-time timer called \textit{rtimer}, which is offered by the Contiki operating system, eliminates timer inaccuracies on the mote side.
The mote uses its light sensor to sense the current brightness.
Once the process is started, an LED-countdown (red $\rightarrow$ blue $\rightarrow$ green) indicates the start of the key initialization procedure.
A \mbox{\textit{2-means}} algorithm is used to create two clusters of \textit{bright} and \textit{dark} values and to compute their mean values~(see \ref{sub:calibration}).
Subsequently a new value is classified as \textit{0} or \textit{1} depending on which mean value is closer to it.
After the successful calibration, synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the failure and the process should be restarted.
A green LED shows that the key material has been received successfully and was saved to the permanent flash memory of the mote.

@@

Subject of debate is also the \textit{Hamming(8,4)-code}.
While it is able to correct one-bit-errors and detect two-bit-errors, it also doubles the amount of data to be sent.
When analyzing the number of corrected bits depending on the transmission rate, the following can be shown:
With long period lengths, the Hamming code does not detect or correct any errors, because the transmission happens without any errors.
When accelerating the transmission, there is a point where many bits are corrected, but the transmission normally succeeds.
If the transmission speed is then set even higher, the whole transmission fails most of the time due to irrecoverable errors.
The borders are 15ms for the \textit{A3} and 63ms for the \textit{S2} (see Figure~\ref{fig:successful_transmissions}).
When we want to achieve the transmission speed of the \textit{A3} with 15ms, we can either use the 15ms and Hamming encoding to correct the bits, or we can halve the amount of data by deactivating the Hamming encoding and transmit with 31ms per bit\footnote{The values are rounded, that is why the half of 31 is 15.}.
So the transmission takes the same time, but actually increases the percentage of successful transmissions.
Errors can still be detected afterwards using the \textit{CRC32} checksum.
Therefore, the \textit{Hamming(8,4)-code} is not useful in that case.

@@

To simplify the initialization of several motes at once, two things need to be done.
First, the Android application should stay in calibration and synchronization phase, until the user presses a button.
Second, the mote should indicate via an LED that it reached the initialization phase~(see \ref{sub:initialization}).
With those two concepts, the key transmission to several motes could work as follows:
The user lays out all motes in a grid next to each other.
Then the phone is placed on top of them and the app is started.
All motes are placed, so that they receive the flickering of the flash light.
The motes are started one after another.
Once the motes are running, the user waits for all of them to indicate via an LED, that they reached the initialization phase.
They are now all waiting for the initialization pattern.
The user presses a button in the application and the phone sends the initialization pattern followed by the data packet.
When the transmission is over, the user can see via a red LED which motes failed to receive the data correctly.
The process needs to be restarted with those motes.
However, this is not implemented and tested yet and could be part of future projects.
